---
title: "Travel Plans"
author: Tingting Zhan
date: "`r Sys.Date()`"
format: 
  html:
    page-layout: full
    html-math-method: katex
number-sections: true
toc: true
toc-location: left
toc-depth: 4
toc-title: ''
editor: source
bibliography: iata.bib
knitr:
  opts_chunk: 
    collapse: true
    comment: "#" 
vignette: >
  %\VignetteIndexEntry{intro}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{=html}
<!--
`r knitr::asis_output('\U1f5dd')`
seems not needed
-->
```

# Introduction

This vignette of R package **`iata`** ([Github](https://github.com/tingtingzhan/iata), [RPubs](https://rpubs.com/tingtingzhan/iata)) documents ...

<!--
-   the creation of `groupedHyperframe` object,
-   the batch processes for a `hyperframe` and/or `groupedHyperframe`, and
-   the aggregation of various statistics over (nested) grouping structure.
-->

R terminology might be different from that of mathematics and statistics. Please refer to @sec-terms for explanation and reference of the terms and abbreviations used in this vignette.

Package **`iata`** `Imports` packages

-   **`cli`** [@cli, v`r packageVersion('cli')`], for attractive command line interfaces
-   `r '\U1f5dd'` **`ggplot2`** [@ggplot2, v`r packageVersion('ggplot2')`], **`scales`** [@scales, v`r packageVersion('scales')`] and **`geomtextpath`** [@geomtextpath, v`r packageVersion('geomtextpath')`], key dependency, for educational visualization
-   `r '\U1f5dd'` **`geosphere`** [@geosphere, v`r packageVersion('geosphere')`], key dependency, for function `geosphere::distGeo()`
-   **`leaflet`** [@leaflet, v`r packageVersion('leaflet')`]
-   **`methods`** shipped with `r R.version$version.string`, for `S4` object oriented system
-   `r '\U1f5dd'` **`plotly`** [@plotly, v`r packageVersion('plotly')`], key dependency, for ..


## Prerequisite

Package **`iata`** requires R version 4.5.0 (released 2025-04-11) or higher ([macOS](https://cran.r-project.org/bin/macosx/), [Windows](https://cran.r-project.org/bin/windows/base/)). An Integrated Development Environment (IDE), e.g., [RStudio](https://posit.co/download/rstudio-desktop/) [@RStudio] or [Positron](https://positron.posit.co/download.html), is not required, but highly recommended. This vignette is created under `r R.version$version.string` using packages **`knitr`** [@knitr, v`r packageVersion('knitr')`], **`quarto`** [@quarto, v`r packageVersion('quarto')` with [Quarto](https://quarto.org/docs/get-started/) v`r quarto::quarto_version()`] and **`rmarkdown`** [@rmarkdown, v`r packageVersion('rmarkdown')`].

```{r}
#| code-fold: true
#| code-summary: "Environment on author's computer"
Sys.info()[c('sysname', 'release', 'machine')]
R.version
```


Experimental (and maybe unstable) features are released *extremely frequently* to [Github](https://github.com/tingtingzhan/iata). [Active developers should use the Github version; suggestions and bug reports are welcome!]{style="background-color: #FFFF00"}

```{r}
#| eval: false
remotes::install_github('tingtingzhan/iata')
```


## Getting Started

Examples in this vignette require that the `search` path has

```{r}
#| message: false
library(iata)
```

```{r}
#| echo: false
op = par(no.readonly = TRUE)
#options(mc.cores = 1L) # for CRAN submission
```


{{< pagebreak >}}

# Flight Path

A mileage run.

```{r}
tripA = 'YYZ-YOW-YFB-YAB-YRB-YGZ, YGZ-YRB-YAB-YFB-YVP-YUL-YYZ' |>
  IATA()
```

```{r}
tripA
```

```{r}
#| fig-height: 4
tripA |> 
  plot()
```


Another mileage run.

```{r}
tripB = 'SFO-HNL-MAJ-KWA-KSA-PNI-TKK-GUM-ROR-MNL-NRT-KIX-HND-SFO' |>
  IATA()
```

```{r}
tripB
```

```{r}
#| fig-height: 4
tripB |> 
  plot()
```


{{< pagebreak >}}

# Airline Status 

```{r}
JSM24 = new(
  Class = 'airfare', 
  carrier = 'AA', 
  depart = 'PHL', arrive = 'PDX', 
  code = 'J', currency = 'usd', 
  basefare = 276.28, tax = 35.82, upgrade = 241.88
)
```

```{r}
JSM24
```



```{r}
#| fig-height: 4
JSM24 |> 
  earnAA()
```


```{r}
#| fig-height: 4
JSM24 |> 
  earnAS()
```

<details><summary>Before Alaska Airlines devaluation 2025</summary>
```{r}
#| fig-height: 4
JSM24 |> 
  earnASold()
```
</details>



```{r}
#| fig-height: 4
JSM24 |> 
  earnBA()
```

<details><summary>Before British Airlines devaluation 2024</summary>
```{r}
#| fig-height: 4
JSM24 |> 
  earnBAold()
```
</details>

# Small Map

```{r}
#| fig-width: 5
'EWR-PHL-JFK-IAD' |> 
  IATA() |> 
  as.leaflet()
```




# Terms & Abbreviations {#sec-terms}

| Term / Abbreviation | Description |
|------------------------------------|------------------------------------|
| CRAN, R | The Comprehensive R Archive Network, <https://cran.r-project.org> |
| `Depends`, `Imports`, `Suggests`, `Enhances` | [*Writing R Extensions*, Section 1.1.3 *Package Dependencies*](https://cran.r-project.org/doc/manuals/r-devel/R-exts.html#Package-Dependencies) |
| [`|>`](https://search.r-project.org/R/refmans/base/html/pipeOp.html) | Forward pipe operator introduced since R version 4.1.0 |
| [`attr`](https://search.r-project.org/R/refmans/base/html/attr.html), [`attributes`](https://search.r-project.org/R/refmans/base/html/attributes.html) | Attributes |
| [`data.frame`](https://search.r-project.org/R/refmans/base/html/data.frame.html) | Data frame |
| `S3`, `generic`, [`methods`](https://search.r-project.org/R/refmans/utils/html/methods.html) | `S3` object oriented system, [`UseMethod`](https://search.r-project.org/R/refmans/base/html/UseMethod.html); [`getS3method`](https://search.r-project.org/R/refmans/utils/html/getS3method.html); <https://adv-r.hadley.nz/s3.html> |
| `S4`, `generic`, `methods` | `S4` object oriented system, [`isS4`](https://search.r-project.org/R/refmans/base/html/isS4.html); [`setClass`](https://search.r-project.org/R/refmans/methods/html/setClass.html); [`setMethod`](https://search.r-project.org/R/refmans/methods/html/setMethod.html); [`getMethod`](https://search.r-project.org/R/refmans/methods/html/getMethod.html); <https://adv-r.hadley.nz/s4.html> |

{{< pagebreak >}}

# References

::: {#refs}
:::
